\section{Introduction}

Many scientific articles confirm the predictive power of quantitative models of biological systems. Kinetic models are some of the most detailed biological models and contain detailed reaction properties that allow predictions of both reaction fluxes and metabolite levels will be, given a set of model parameters. An early example of a kinetic model is that of the Teusink \emph{et al.} \cite{teusink2000} kinetic model of yeast glycolysis, which with updating of branching reactions was able to predict the \emph{in vivo} flux of half the model reactions to within a factor of 2. In comparison, stoichiometric models of biological systems don't contain enzyme kinetic parameters but instead focus solely on which metabolites participate in which reactions. The iND750 version of the \emph{S.cerevisiae} genome scale model \cite{duarte2004} contains 750 reactions, and the growth rate of the model is simulated by maximising a given objective reaction such as the production of biomass. The growth rate is then defined in the combination of internal reaction fluxes.

Regardless of which type of model used, a common objective is then to analyse, simulate and test the model to observe what predictions the model can make about the corresponding real-life biological system. One avenue of analysis of kinetic biological model is metabolic control analysis, or MCA \cite{fell1997}. One of the key points of MCA is there is not single rate limiting step in a set of reactions, but instead each reaction and metabolite in a system may share a portion of control over the other reaction rates and metabolite levels in the system. As an example the amount of control an enzyme has over a pathway can be estimated by measuring the flux through the system, then reducing the activity of the same enzyme and examining the difference in system flux. This is described in Eq. \ref{fcc}, where $J$ and $Block$ are the fluxes at reaction $m$ and system $i$ respectively. The sign $C_{Block_i}^{J_m}$ is the flux control coefficient, a measure of how sensitive the pathway is to changes in flux through the reaction, $J_{m}$. The flux control coefficients will sum to 1 over a linear pathway as shown in Eq. \ref{summation_theorum}, the summation theorem.

\begin{equation}\label{fcc}
C_{Block_i}^{J_m} = \frac{dJ_m}{J_m}\div\frac{dBlock_i}{Block_i}
\end{equation}

\begin{equation}\label{summation_theorum}
\sum_{i=1}^{n}C_{Block_i}^{J_m} = 1
\end{equation}

An example of how metabolic control analysis can be successfully applied to real biological systems is the examination of tryptophan biosynthesis carried out by Niederberger \emph{et al.} \cite{Niederberger1992}. Using a five-enzyme pathway model involved in the production of tryptophan, it was shown that increasing/decreasing levels of individual enzymes has very little effect on the overall flux through the pathway. This was explained as an effect of the distribution of flux control over the length of the pathway, with no particular enzyme having a large flux control coefficient. However, increasing the concentration of all five enzymes, and therefore the flux through their catalysed reactions, resulted in a synergistic increase in tryptophan flux - much greater than would be expected if the increased enzyme fluxes were only cumulative.

The number of enzyme parameter needed to produce a kinetic model mean that the number of reactions included can be a magnitude smaller than that of genome scale models. The small number of reactions therefore means that the estimation of  control coefficients are limited to the small number of available reactions. For instance, at the time of writing this thesis, there exists no model where it would be possible to derive control coefficients for reactions across different parts of the cell metabolism such as both energy metabolism and amino acid. Existing genome scale models contain reactions contain the majority of the central metabolism for the organism they describe, which solves the problem of limited model size. The restriction of genome scale models is that they do not contain enzyme kinetic data, and so it is not possible to estimate metabolic control coefficients in the same way as kinetic models.

This section highlights the methods and results used in this analysis for trying to estimate the importance and sensitivity of the parts of the genome scale models to simulated perturbations in their activity.
